### 2019-05-05 pages/math/semi-comb/project/paper.tex
Update

\documentclass{amsart}
\usepackage[T1]{fontenc}
\usepackage{mathpazo}
%\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
%\usepackage{rotating}
%\usepackage{color}
%\usepackage{tabu}
%\usepackage{parskip}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{calc}
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{calc}
%\usepackage{xcolor-solarized}
%\usepackage{minted}
%\usepackage{MnSymbol}
%\usepackage[libertine]{newtxmath}
%\newcommand*\circled[1]{\tikz[baseline= (char.base)]{
    %\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%\usepackage{xypic}
%\usepackage[tightpage]{preview} % display
% Math operators
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\F}{\mathbb{F}}
\DeclareMathOperator{\sge}{\mathsf{e}}
\DeclareMathOperator{\sgm}{\mathsf{m}}
\DeclareMathOperator{\sgG}{\mathsf{G}}
\DeclareMathOperator{\sgg}{\mathsf{g}}
\DeclareMathOperator{\sgF}{\mathsf{F}}
\DeclareMathOperator{\sgt}{\mathsf{t}}
\DeclareMathOperator{\sgL}{\mathsf{L}}
\DeclareMathOperator{\sgZ}{\mathsf{Z}}
\DeclareMathOperator{\sgAp}{Ap}
\DeclareMathOperator{\sgPF}{PF}
\DeclareMathOperator{\sgconv}{conv}
\DeclareMathOperator{\sgsupp}{supp}
\DeclareMathOperator{\calO}{\mathcal{O}}
\DeclareMathOperator{\Grho}{\prescript{}{\rho}G}
% Theorem environments
\theoremstyle{plain}
\newtheorem*{thm}{Theorem}
\newtheorem*{lemma}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{cor}{Corollary}
\newtheorem*{prop}{Proposition}
\newtheorem*{conj}{Conjecture}
\theoremstyle{remark}
\newtheorem*{example}{Example}
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
% Paired math operators
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
% Math helper functions
\newcommand{\congr}[2]{{[#1]}_{#2}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\genset}[1]{{\langle{}#1\rangle{}}}
\newcommand{\abs}[1]{\lvert{}#1\rvert{}}
\newcommand{\norm}[1]{\lVert{}#1\rVert{}}
\newcommand{\bmx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand{\pmx}[1]{\begin{pmatrix}#1\end{pmatrix}}
  \newcommand{\sbmx}[1]{{\left[\begin{smallmatrix}#1\end{smallmatrix}\right]}}
\newcommand{\smallxymatrix}[2]{\xymatrix@C-=#1cm@R-=#1cm{#2}}
\renewcommand\labelitemi{\small$\bullet$}
\begin{document}
%
%
%
\noindent
\begin{quote}
  (13) \emph{Flow polytopes.}
  A \emph{directed graph} is a collection of dots (called \emph{vertices}) and
  arrow between them (called \emph{edges}). A \emph{flow} on a directed graph is a
  way to label each edge of $G$ by an element of $\Z$ (or $\Z_n$) so that at each
  vertex $v$, the sum of the edges entering $v$ equals the sum of the edges
  leaving $v$ (i.e.\ the in \emph{flow} at each vertex equals the \emph{out
  flow}). The flows on a given directed graph can be counted using a certain
  family of polytopes, whose integer points each coincide with a flow on $G$.
  \\[1em]
  \noindent
  Source:
  \emph{Combinatorial reciprocity theorems} (M. Beck, R. Sanyal), Chapters 1 and 7.
\end{quote}

@@

% 1.2 Flows on Graphs notes

A graph $G=(V,E)$ is a collection of $V$ a set of vertices and $E$ a set of
edges between these vertices. Now, a graph doesn't have to have any vertices or
any edges (except that if it has edges there must be vertices to which those edges
connect).
Also note that the vertices of a graph can be elements of some arbitrary $\F^n$
space, but for this report we'll only concern ourselves with graphs in $\Z^2$
Euclidean space.
\begin{definition}[Component, connected component]
  A \emph{connected component} (or simply a \emph{component}) of a graph $G$ is
  a subgraph (a subset of the vertices and edges of $G$) in which any two
  vertices are either directed connect via an edge or are connected to
  one-another via a \emph{path} (a series of vertices and edges). A single
  vertex with no connecting edges is in itself a connected component.

@@

Any graph then is the disjoint union of connected components.
For example figure~\ref{fig:disjoint-union} can be called a single graph that is the disjoint
union of three distinct connected components.
\begin{figure}[h]
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (.5,.5) node (a) {}
      (0,0) node (b) {} (1,0) node (c) {} (1,1) node (d) {} (0,1) node (e) {}
      (2,1) node (f) {} (2,0) node (g) {} (3,0) node (h) {};
    \draw (a);
    \draw (b) -- (c) -- (d) -- (e) -- (b);
    \draw (f) -- (g) -- (h);
  \caption[Disjoin union]{}
  \label{fig:disjoint-union}
\begin{lemma}
  A connected component of the graph $G$ is a maximal subgraph of $G$ in which
  any two nodes are connect by a path.
\end{lemma}
\begin{lemma}
  A graph $G$ is \emph{connected} if it has only one connected component.
\end{lemma}
\begin{definition}[Bridge]
  A \emph{bridge} is a single edge which serves as a connection between two connected
  components which if removed would increase the number of connected
  components of the graph by one.
For example, removing the bridge in figure~\ref{fig:bridge} turns a single
connected component into two triangular connected components:
one single connected graph into a graph with two connected subgraphs.
\begin{figure}[h]
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
    \draw[dashed] (a) -- (d);
  \hspace{1cm}
  \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
  \caption[Bridges]{A graph with and without a bridge.}
  \label{fig:bridge}
So far we've only looked at graph which are \emph{unoriented}.
\begin{definition}[Directed, oriented graph]
  A \emph{directed} or \emph{oriented graph} is a graph is a graph on which we
  define an \emph{orientation} via a subset $\rho$ of the edges $E$ of the
  graph. For each edge $e=v_i v_j\in E$ (connecting the vertices $v_i,v_j\in
  V$) with $i<j$ we direct:
  \[
    v_i\stackrel{e}{\leftarrow}v_j\text{ if }e\in\rho\text{, and }
    v_i\stackrel{e}{\rightarrow}v_j\text{ if }e\notin\rho.
  \]
This definition may be confusing at first glance because $\rho$
may be missing some of the edges of $E$. What this definition attempts to succinctly
encapsulate is only the edges with their direction pointing from the vertex of higher
index to the vertex of lower index (which does certainly feel backwards),
with all the remaining edges not in $\rho$ directed in the opposite direction,
from lower to higher index.
Figure~\ref{fig:directed-graph} tries to illustrate this encapsulation for a graph
with five vertices and an orientation $\rho=\{14,23,24\}$.
\begin{figure}[h]
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style={draw,shape=circle,fill=white,inner sep=1}]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw (v_4) -- (v_3) -- (v_2) -- (v_1) -- (v_4) -- (v_2);
    \caption{The undirected graph.}
  %\hspace{1cm}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering

@@

        every node/.style={draw,shape=circle,fill=white,inner sep=1},

@@

      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[postaction={decorate}] (v_4) -- (v_1);
      \draw[postaction={decorate}] (v_4) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_2);
      \draw (v_3) -- (v_4);
      \draw (v_1) -- (v_2);

@@

    \caption{$\rho$ edges marked.}

@@

  \begin{subfigure}[t]{0.3\textwidth}
    \centering
        every node/.style={draw,shape=circle,fill=white,inner sep=1},
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[postaction={decorate}] (v_4) -- (v_1);
      \draw[postaction={decorate}] (v_4) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_2);
      \draw[postaction={decorate}] (v_3) -- (v_4);
      \draw[postaction={decorate}] (v_1) -- (v_2);
    \caption{$E\setminus\rho$ edges marked.}
  \caption[A directed graph]{A directed graph.}
  \label{fig:directed-graph}

>>>>

\documentclass{amsart}
\usepackage[T1]{fontenc}
%\usepackage{mathpazo}
%\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
%\usepackage{rotating}
%\usepackage{color}
%\usepackage{tabu}
%\usepackage{parskip}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{calc}
\usepackage{xfrac}
\usepackage{mathtools}
\usepackage{cancel}
\usepackage{framed}
\usepackage{nth}
\usepackage{tikz}
\usetikzlibrary{calc,decorations.markings}
%\usepackage{xcolor-solarized}
%\usepackage{minted}
%\usepackage{MnSymbol}
%\usepackage[libertine]{newtxmath}
%\newcommand*\circled[1]{\tikz[baseline= (char.base)]{
    %\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%\usepackage{xypic}
%\usepackage[tightpage]{preview} % display
% Math operators
\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\N}{\mathbb{N}}
\DeclareMathOperator{\R}{\mathbb{R}}
\DeclareMathOperator{\C}{\mathbb{C}}
\DeclareMathOperator{\F}{\mathbb{F}}
\DeclareMathOperator{\sge}{\mathsf{e}}
\DeclareMathOperator{\sgm}{\mathsf{m}}
\DeclareMathOperator{\sgG}{\mathsf{G}}
\DeclareMathOperator{\sgg}{\mathsf{g}}
\DeclareMathOperator{\sgF}{\mathsf{F}}
\DeclareMathOperator{\sgt}{\mathsf{t}}
\DeclareMathOperator{\sgL}{\mathsf{L}}
\DeclareMathOperator{\sgZ}{\mathsf{Z}}
\DeclareMathOperator{\sgAp}{Ap}
\DeclareMathOperator{\sgPF}{PF}
\DeclareMathOperator{\sgconv}{conv}
\DeclareMathOperator{\sgsupp}{supp}
\DeclareMathOperator{\calC}{\mathcal{C}}
\DeclareMathOperator{\calF}{\mathcal{F}}
\DeclareMathOperator{\calO}{\mathcal{O}}
\DeclareMathOperator{\Grho}{\prescript{}{\rho}G}
% Theorem environments
\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem*{thm*}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{cor}{Corollary}
\newtheorem*{prop}{Proposition}
\newtheorem*{conj}{Conjecture}
\theoremstyle{remark}
\newtheorem*{example*}{Example}
\theoremstyle{definition}
\newtheorem*{definition*}{Definition}
% Special theorem displays
\newenvironment{example}%
	{\begin{leftbar}\begin{example*}
}{%
	\end{example*}\end{leftbar}
}
\newenvironment{definition}%
	{\begin{leftbar}\begin{definition*}
}{%
	\end{definition*}\end{leftbar}
}
% Paired math operators
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
% Math helper functions
\newcommand{\congr}[2]{{[#1]}_{#2}}
\newcommand{\set}[1]{\{#1\}}
\newcommand{\genset}[1]{{\langle{}#1\rangle{}}}
\newcommand{\abs}[1]{\lvert{}#1\rvert{}}
\newcommand{\norm}[1]{\lVert{}#1\rVert{}}
\newcommand{\bmx}[1]{\begin{bmatrix}#1\end{bmatrix}}
\newcommand{\pmx}[1]{\begin{pmatrix}#1\end{pmatrix}}
  \newcommand{\sbmx}[1]{{\left[\begin{smallmatrix}#1\end{smallmatrix}\right]}}
\newcommand{\smallxymatrix}[2]{\xymatrix@C-=#1in@R-=#1in{#2}}
\renewcommand\labelitemi{\small$\bullet$}
%
%\makeatletter
\tikzset{%
  gdot/.style={draw,shape=circle,fill=black,inner sep=1},
  ddot/.style={draw,shape=circle,fill=white,inner sep=1},
  dedge/.style={draw=black!40},
  midarrow/.style={postaction={decorate},decoration={markings,mark=at position 0.5 with {\arrow{>}}}}
}
%\makeatother
\begin{document}
%
%
%

\section*{Prompt:}

\noindent
\begin{quote}
  (13) \emph{Flow polytopes.}
  A \emph{directed graph} is a collection of dots (called \emph{vertices}) and
  arrow between them (called \emph{edges}). A \emph{flow} on a directed graph is a
  way to label each edge of $G$ by an element of $\Z$ (or $\Z_n$) so that at each
  vertex $v$, the sum of the edges entering $v$ equals the sum of the edges
  leaving $v$ (i.e.\ the in \emph{flow} at each vertex equals the \emph{out
  flow}). The flows on a given directed graph can be counted using a certain
  family of polytopes, whose integer points each coincide with a flow on $G$.
  \\[1em]
  \noindent
  Source:
  \emph{Combinatorial reciprocity theorems} (M. Beck, R. Sanyal), Chapters 1 and 7.
\end{quote}

@@

Graphs are a major area of concern in mathematics, having an entire field to
study dedicated to them. Many different families of graphs exist, but for this
paper we'll concern ourselves with directed graphs and $\Z_n$-flows on these
graphs, all concepts which the first section introduces. The second section
is dedicated to a reciprocity expression for counting the number of possible
flows on directed graphs, and to developing some insight into it and its proof.

% ------------------------------------------------------------------------------
% 1.2
% ------------------------------------------------------------------------------

\section{Introduction to graphs and flows}

\subsection{Graphs}

A graph $G=(V,E)$ is a collection of $V$, a set of vertices, and $E$, a set of
edges between the vertices.
Now, a graph doesn't have to have any vertices, nor
does it have to to have any edges (except that if it has edges there must be
vertices to which those edges connect).
Also note that the nodes of a graph can be elements of some arbitrary
d-dimensional space, but for this report we'll only concern ourselves with
graphs in some two-dimensional space.
(Note that the word ``node'' is analogous to ``vertex'' and that this paper
will do plenty of haphazard switching between the two of them.)
\begin{definition}[Component, connected component]
  A \emph{connected component} (or simply a \emph{component}) of a graph $G$ is
  a subgraph (a subset of the vertices and edges of $G$) in which any two
  vertices are either directly connected via an edge or are connected to
  one-another via a \emph{path} (a series of vertices and edges). A single
  vertex with no connecting edges is in itself a connected component.
\end{definition}
Any graph then is the disjoint union of connected components.
For example Figure~\ref{fig:disjoint-union} can be called a single graph that is the disjoint
union of three distinct connected components.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (.5,.5) node (a) {}
      (0,0) node (b) {} (1,0) node (c) {} (1,1) node (d) {} (0,1) node (e) {}
      (2,1) node (f) {} (2,0) node (g) {} (3,0) node (h) {};
    \draw (a);
    \draw (b) -- (c) -- (d) -- (e) -- (b);
    \draw (f) -- (g) -- (h);
  \end{tikzpicture}
  \caption[Disjoin union]{}
  \label{fig:disjoint-union}
\end{figure}
\begin{lemma*}
  A connected component of the graph $G$ is a maximal subgraph of $G$ in which
  any two nodes are connect by a path.
\end{lemma*}
\begin{lemma*}
  A graph $G$ is \emph{connected} if it has only one connected component.
\end{lemma*}
\begin{definition}[Bridge]
  A \emph{bridge} is a single edge which serves as a connection between two connected
  components which, if removed, would increase the number of connected
  components of the graph by one.
\end{definition}
For example, removing the bridge in Figure~\ref{fig:bridge} turns a single
connected component into two triangular connected components:
one complete graph into a graph with two complete subgraphs.
\begin{figure}[ht]
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
    \draw[dashed] (a) -- (d);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}[every node/.style=gdot]
    \path
      (-.5,.5) node (a) {} ($({-(1+sqrt(3))/2},0)$) node (b) {} ($({-(1+sqrt(3))/2},1)$) node (c) {}
      ( .5,.5) node (d) {} ($({ (1+sqrt(3))/2},0)$) node (e) {} ($({ (1+sqrt(3))/2},1)$) node (f) {};
    \draw (a) -- (b) -- (c) -- (a);
    \draw (d) -- (e) -- (f) -- (d);
  \end{tikzpicture}
  \caption[Bridges]{A graph with and without a bridge.}
  \label{fig:bridge}
\end{figure}

\subsection{Directed graphs}

So far we've only looked at \emph{unoriented} graphs, but
now we introduce a notion of direction, or orientation, to graphs.
\begin{definition}[Directed, oriented graph]
  A \emph{directed} or \emph{oriented graph} is a graph on which we
  define an \emph{orientation} via a subset $\rho$ of the edges $E$ of the
  graph. For each edge $e=v_i v_j\in E$ (connecting the vertices $v_i,v_j\in
  V$) with $i<j$ we direct:
  \[
    v_i\stackrel{e}{\leftarrow}v_j\text{ if }e\in\rho\text{, and }
    v_i\stackrel{e}{\rightarrow}v_j\text{ if }e\notin\rho.
  \]
\end{definition}
This definition may be confusing at first glance because $\rho$
may be missing some of the edges of $E$. What this definition attempts to succinctly
encapsulate is only the edges with their direction pointing from the vertex of higher
index to the vertex of lower index (which does certainly feel backwards),
with all the remaining edges not in $\rho$ directed in the opposite direction,
from lower to higher index.
Figure~\ref{fig:directed-graph} tries to illustrate this encapsulation for a graph
with five vertices and an orientation $\rho=\{14,23,24\}$.
\begin{figure}[ht]
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw (v_4) -- (v_3) -- (v_2) -- (v_1) -- (v_4) -- (v_2);
    \end{tikzpicture}
    \caption{The undirected graph.}
  \end{subfigure}
  %\hspace{1in}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style=gdot]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw (v_3) -- (v_4);
      \draw (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$\rho$ edges marked.}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \begin{tikzpicture}[every node/.style={gdot}]
      \path
        (0,1) node (v_1)[label=left:$v_1$]{}
        (1,1) node (v_2)[label=right:$v_2$]{}
        (1,0) node (v_3)[label=right:$v_3$]{}
        (0,0) node (v_4)[label=left:$v_4$]{};
      \draw[midarrow] (v_4) -- (v_1);
      \draw[midarrow] (v_4) -- (v_2);
      \draw[midarrow] (v_3) -- (v_2);
      \draw[midarrow] (v_3) -- (v_4);
      \draw[midarrow] (v_1) -- (v_2);
    \end{tikzpicture}
    \caption{$E\setminus\rho$ edges marked.}
  \end{subfigure}
  \caption[A directed graph]{A directed graph.}
  \label{fig:directed-graph}
\end{figure}

As an aside, directed graphs are the subject of many algorithms in computer
science, often pertaining to finding optimal ways to traverse the paths of said
graphs. In some models each edge is considered to be of equal ``weight'' as any
other. That is if two edges originating from the same node both point to the
same node, there should be no real preference to traveling down one edge or the
other to get to the destination. On the other hand we might assign a weight to
each edge, presenting a cost-benefit analysis in choosing a path to take (as in
Figure~\ref{fig:path-finding} where the edge with weight 5 now becomes most
expensive to traverse versus the one with weight 2).
\begin{figure}[ht]
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$\phantom 5$} (b);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}
    \path (0,0) node[gdot] (a){} (2,0) node[gdot] (b){};
    \draw[midarrow] (a) to [out=45, in=135] node[above]{$2$} (b);
    \draw[midarrow] (a) to [out=315,in=225] node[below]{$5$} (b);
  \end{tikzpicture}
  \caption{A directed graph with and without weights.}
  \label{fig:path-finding}
\end{figure}
Intuitively these weights are often simple $\Z$ integers.

\subsection{Flows}

Moving on we introduce \emph{flow} values to edges; similar
values to weights, but living not within $\Z$ but instead within an Abelian
group $\Z_n=\sfrac\Z n$.
%(There may also be optimal path algorithms associated
%with flow values, similar to weights on directed edges, but for this paper our
%focus is elsewhere.)
\begin{definition}[$\Z_n$-flow]
  A \emph{$\Z_n$-flow} is a map $f:E\to\Z_n$ which to each $e$ edge of $E$
  assigns a flow ${f(e)\in\Z_n}$ such that there is ``conservation of flow'' at
  every $v\in V$ vertex, by which me mean the sum of the flows into a vertex $v$
  is equivalent to the sum of the flows out it:
  \[
    \sum_{\stackrel e \to v}f(e)=\sum_{v\stackrel e \to}f(e).
  \]
\end{definition}
(And it's important to note equivalence here since $\sum f(e)\in\Z_n$.)
These functions and more specifically counting them will be the ultimate
purpose of this paper.
\begin{definition}[$f$ support]
  The \emph{support} of $f$, denoted $\sgsupp(f)$, is the subset of edges $e$
  whose flows are non-zero:
  \[
    \sgsupp(f)=\{e\in E:f(e)\ne 0\}\subseteq E
  \] 
\end{definition}
If $\sgsupp(f)=E$ (all $e$ edges have $f(e)\ne0$), then $f$ is considered
``nowhere-zero''. For the rest of this report we'll be concerned with these
nowhere-zero $\Z_n$-flows, in particular with counting how many of them can be
made for any particular graph and Abelian group.
%
Figure~\ref{fig:graph-flows} illustrates two different but satisfactory
$\Z_5$-flows on a graph. Without the context of the Abelian group $Z_5$
these flows wouldn't seem to work, but, since $3+2\equiv 0\bmod 5$,
the top and bottom left vertices both have conservation of flow
equal to zero.
\begin{figure}[ht]
  \begin{tikzpicture}
    \path
      (0,1) node (a)[gdot]{}
      ($({cos(  -pi/6 r)},{sin(  -pi/6 r)})$) node (b)[gdot]{}
      ($({cos(-5*pi/6 r)},{sin(-5*pi/6 r)})$) node (c)[gdot]{};
    \draw[midarrow] (a) -- node[right]{0} (b);
    \draw[midarrow] (a) -- node[left]{0} (c);
    \draw[midarrow] (b) -- node[below]{0} (c);
  \end{tikzpicture}
  \hspace{1in}
  \begin{tikzpicture}
    \path
      (0,1) node (a)[gdot]{}
      ($({cos(  -pi/6 r)},{sin(  -pi/6 r)})$) node (b)[gdot]{}
      ($({cos(-5*pi/6 r)},{sin(-5*pi/6 r)})$) node (c)[gdot]{};
    \draw[midarrow] (a) -- node[right]{2} (b);
    \draw[midarrow] (a) -- node[left]{3} (c);
    \draw[midarrow] (b) -- node[below]{2} (c);
  \end{tikzpicture}
  \caption{An everywhere-zero $\Z_5$-flow and a nowhere-zero $\Z_5$-flow.}
  \label{fig:graph-flows}
\end{figure}

Going forward, $G$ will always refer to a 2-dimensional graph with $V$ vertices
and $E$ edges, $\rho$ to an arbitrary (but fixed) orientation of $G$
and $\Z_n$ to an Abelian group to which $f$ a $\Z_n$-flow maps.
\begin{definition}
  Define a counting function:
  \[
    \varphi_G(n)=\text{the number of $f$ nowhere-zero $\Z_n$-flows on $G$}
  \] 
\end{definition}
As it happens\ldots
\begin{prop}
  The flow-counting function $\varphi_G(n)$ is independent of the orientation
  $\rho$ of $G$.
\end{prop}
(The proof of this is a ``left to the reader'' exercise in the original paper.)
\begin{prop}
  $G$ will not have any nowhere-zero flow if $G$ has a \emph{bridge}---an edge
  whose removal increases the number of connected components of $G$.
\end{prop}

\subsection{Dual graphs}

Consider a graph $G$ as subdividing the plane into connected regions. If two
points lie within a region then they can be connected without intersecting an
edge of $G$. Regions which are neighbors have an edge of $G$ which separate them
(their topological closures sharing a proper 1-dimensional part of their
boundaries).
\begin{definition}[Dual graph]
  The \emph{dual graph} of a graph $G=(V,E)$ is the graph $G^*=(V^*,E^*)$ with
  vertices corresponding to the regions of $G$.
  %
  Two vertices $u^*,v^*\in V^*$
  which correspond to two regions $R_{u^*}$ and $R_{v^*}$ of $G$ are connected via edge $e^*\in
  E^*$ if an original edge $e\in E$ is properly contained within the boundaries
  of $R_{u^*}$ and $R_{v^*}$.

@@

%Figure~\ref{fig:dual-incomplete} shows a graph with four vertices (solid)
%along with the vertices (hollow) of its dual to demonstrate the regions which
%$G$ subdivides the plane into.
%(However the edges are omitted.
%See Figure~\ref{fig:graphs-and-duals} for complete illustrations of graphs and
%their duals.)
%\begin{figure}[ht]
    %\begin{tikzpicture}
      %\path
        %(0,0) node (a)[gdot]{}
        %(1,0) node (b)[gdot]{}
        %(1,1) node (c)[gdot]{}
        %(0,1) node (d)[gdot]{};
      %\draw (a) -- (b) -- (c) -- (d) -- (a);
      %\path (.5,.5) node[ddot]{} (1.5,.5) node[ddot]{};
  %\end{tikzpicture}
  %\caption{A graph and its (incompletely portrayed) dual graph.}
  %\label{fig:dual-incomplete}
%\end{figure}
In other words there is one $e^*$ for
every edge $e$ that lies between two regions of $G$; it connects the $G^*$
vertices corresponding to the two regions; and it crosses over its
corresponding $e$ edge ($e$ and $e^*$ intersect).
%
But as a single illustration says a thousand words,
Figure~\ref{fig:graphs-and-duals} illustrates a few examples of graphs
(solid dot vertices) and their duals (hollow dot vertices).
\begin{figure}[ht]
  \begin{subfigure}[t]{.3\textwidth}

@@

        gdot/.style={draw,shape=circle,fill=black,inner sep=1},
        ddot/.style={draw,shape=circle,fill=white,inner sep=1},

@@

      \begin{scope}
        %\clip (-2,-1) rectangle (2,2);
        \draw (0,0) node[gdot](a){} -- (1,0) node[gdot](b){};
        \node[ddot](c) at (.5,-.5) {};
        \draw[dedge] (c) to [out=180,in=90,looseness=60] (c);
      \end{scope}

@@

    \caption{Simplest graph with a bridge.}

@@

  %\begin{subfigure}[t]{.3\textwidth}
    %\begin{tikzpicture}[
        %gdot/.style={draw,shape=circle,fill=black,inner sep=1},
        %ddot/.style={draw,shape=circle,fill=white,inner sep=1},
        %decoration={markings,mark=at position 0.5 with {\arrow{>}}}
      %]
      %\begin{scope}
        %%\clip (-2,-1) rectangle (2,2);
        %\draw (0,0) node[gdot]{} -- (1,0) node[gdot]{} -- ($(1/2,{sqrt(3)/2})$) node [gdot]{} -- cycle;
        %\node[ddot] (a) at ($(1/2,{-sqrt(2)/3})$) {};
        %\node[ddot] (b) at ($(1/2,{sqrt(3)/6})$) {};
        %\draw[dedge] (a) -- (b);
        %\draw[dedge] (a) to [out=180,in=135,looseness=4] (b);
        %\draw[dedge] (a) to [out=0,in=45,looseness=4] (b);
      %\end{scope}
    %\end{tikzpicture}
    %\caption{A complete graph.}
  %\end{subfigure}
  %\begin{subfigure}[t]{.3\textwidth}
    %\begin{tikzpicture}[
      %gdot/.style={draw,shape=circle,fill=black,inner sep=1},
      %ddot/.style={draw,shape=circle,fill=white,inner sep=1},
      %decoration={markings,mark=at position 0.5 with {\arrow{>}}}
    %]
    %\begin{scope}
      %%\clip (-2,-1) rectangle (2,2);
      %\draw (0,0) node[gdot]{} -- (0,1) node[gdot]{};
      %\draw (0,0) node[gdot]{} -- ($({sqrt(3)/2},1/2)$) node[gdot]{} -- (0,1) node[gdot]{} -- ($({-sqrt(3)/2},1/2)$) node[gdot]{} -- cycle;
      %\node[ddot] (a) at (0,-1/2){};
      %\node[ddot] (b) at ($({-sqrt(3)/6},1/2)$){};
      %\node[ddot] (c) at ($({ sqrt(3)/6},1/2)$){};
      %\draw[dedge] (a) to [out=120,in=240] (b);
      %\draw[dedge] (a) to [out=180,in=120,looseness=4] (b);
      %\draw[dedge] (a) to [out=60,in=300] (c);
      %\draw[dedge] (a) to [out=0,in=60,looseness=4] (c);
      %\draw[dedge] (b) -- (c);
    %\end{scope}
    %\end{tikzpicture}
    %\caption{Graph with two connected components.}
  %\end{subfigure}
  \caption{Various graphs and their duals.}
  \label{fig:graphs-and-duals}
